Verifying Fine-Grained and Lock-Free Binary Search Trees

### 2022-02-15 concurrency/Verification of Concurrent Binary Search Tree.tex
Updated new diagrams

\subsubsection{Key Ranges}
When we traverse the BST looking for a key $k$, we will reach either a node containing $k$, or an empty node where $k$ would appear if it was in the tree. To prove correctness of this procedure, we need ghost state that tracks where each key ``should appear'' in the current tree. As observed by \citet{krishna2017flow}, this can be done by associating each node with a lower and upper bound on the keys appearing in the subtree rooted at that node. We refer to the product of these bounds as a \emph{range}. The range of a node is inherited from its parent, based on the parent's key: if node $n$ has range $(l, r)$ and key $k$, then its left child has range $(l, k)$ and its right child has range $(k, r)$. Figure~\ref{range_bst} shows an example of a BST with each node labeled with its range, starting with $(-\infty, +\infty)$ at the root and propagated to the empty \emph{leaf} nodes. 
\begin{figure}[htb]
\centering
\includegraphics[width=70mm,scale=0.5]{FIG/range_prop.png}
\caption{An example of BST with each node labeled with \emph{range}; lower and upper bound for a key}
\label{range_bst}
\end{figure}
At the leaves, these ranges partition the space of all possible keys. If we reach a leaf node with range $(a, b)$, then we know that keys between $a$ and $b$ 1) are not currently in the tree and 2) can correctly be inserted at this leaf.  For example, suppose we want to insert a key $38$ into the tree in Figure~\ref{range_bst}. The right child of the node with key 35 is a leaf with range $(35, 40)$, so it is guaranteed to be the right place to insert 38.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=150mm,scale=0.5]{FIG/extract_insert.png}
\caption{A visual depiction of the change in global state of BST during $\mathsf{insert(k,v)}$ operation }
\label{extract_insert}
\end{figure}
\begin{figure}[htb]
\centering
\includegraphics[width=150mm,scale=0.5]{FIG/extract_insert_2.png}
\caption{A visual depiction of the change in global state of BST during $\mathsf{insert(k,v)}$ operation, where key $k$ already exists in the tree $t$ }
\label{extract_insert2}
\end{figure}

@@

\usetikzlibrary{positioning}

\forestset{
   lab/.style={
        label={[anchor=west, scale=0.75] north east:#1},

>>>>

\subsubsection{Key Ranges}
When we traverse the BST looking for a key $k$, we will reach either a node containing $k$, or an empty node where $k$ would appear if it was in the tree. To prove correctness of this procedure, we need ghost state that tracks where each key ``should appear'' in the current tree. As observed by \citet{krishna2017flow}, this can be done by associating each node with a lower and upper bound on the keys appearing in the subtree rooted at that node. We refer to the product of these bounds as a \emph{range}. The range of a node is inherited from its parent, based on the parent's key: if node $n$ has range $(l, r)$ and key $k$, then its left child has range $(l, k)$ and its right child has range $(k, r)$. Figure~\ref{range_bst} shows an example of a BST with each node labeled with its range, starting with $(-\infty, +\infty)$ at the root and propagated to the empty \emph{leaf} nodes.

\begin{figure}[htb]
\centering
\includegraphics[width=70mm,scale=0.5]{FIG/range_prop.png}
\caption{An example of BST with each node labeled with \emph{range}; lower and upper bound for a key}
\label{range_bst}
\end{figure}

\usetikzlibrary{positioning}
\forestset{
   lab/.style={
        label={[anchor=west, scale=0.75] north east:#1},
    },
}
\begin{forest}
for tree={
    grow=south,
    circle, draw, minimum size=3ex, inner sep=1pt,
    s sep=3mm
        }
    % my label/.style={
        % label={[anchor=south west] north east:#1},
    % },
    [ 20, tikz={\node[right=0pt of .north east, scale=0.75]  {$(-\infty,\infty)$};}
        [15, tikz={\node[left=0pt of .north west, scale=0.75]  {$(-\infty,20)$};}]
        [30, tikz={\node[right=0pt of .north east, scale=0.75]  {$(20,\infty)$};}
            [25, tikz={\node[left=0pt of .north west, scale=0.75]  {$(20,30)$};}]
            [50, tikz={\node[right=0pt of .north east, scale=0.75]  {$(30,\infty)$};}
                [45, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,50)$};}
                    [40, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,45)$};}
                        [35, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,40)$};}
                            [, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,35)$};}]
                            [, tikz={\node[right=0pt of .north east, scale=0.75]  {$(35,40)$};}]
                        ]
                        [,no edge, draw=none]
                    ]
                    [,no edge, draw=none]
                ]
                [,no edge, draw=none]
            ]
        ]
    ]
\end{forest}

At the leaves, these ranges partition the space of all possible keys. If we reach a leaf node with range $(a, b)$, then we know that keys between $a$ and $b$ 1) are not currently in the tree and 2) can correctly be inserted at this leaf.  For example, suppose we want to insert a key $38$ into the tree in Figure~\ref{range_bst}. The right child of the node with key 35 is a leaf with range $(35, 40)$, so it is guaranteed to be the right place to insert 38.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=150mm,scale=0.5]{FIG/extract_insert.png}
\caption{A visual depiction of the change in global state of BST during $\mathsf{insert(k,v)}$ operation }
\label{extract_insert}
\end{figure}

\begin{forest}
for tree={
    grow=south,
    circle, draw, minimum size=3ex, inner sep=1pt,
    s sep=3mm
        }
    [, phantom, for children={fit=band}, s sep+=20mm,
         before drawing tree={%
          tikz+={%
            \node (a) [inner sep=0pt, fit=(!1) (!1L) (!1F)] {};
            \node (b) [inner sep=0pt, fit=(!l) (!lL) (!lF)] {};
            \node [anchor=center] at ($(a.east)!1/2!(b.west)$) {$\longrightarrow\quad$};
          },
        },
        [ 20, tikz={\node[right=0pt of .north east, scale=0.75]  {$(-\infty,\infty)$};}
            [15, tikz={\node[left=0pt of .north west, scale=0.75]  {$(-\infty,20)$};}]
            [30, tikz={\node[right=0pt of .north east, scale=0.75]  {$(20,\infty)$};}
                [25, tikz={\node[left=0pt of .north west, scale=0.75]  {$(20,30)$};}]
                [50, tikz={\node[right=0pt of .north east, scale=0.75]  {$(30,\infty)$};}
                    [45, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,50)$};}
                        [40, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,45)$};}
                            [35, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,40)$};}
                                [, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,35)$};}]
                                [, tikz={\node[right=0pt of .north east, scale=0.75]  {$(35,40)$}; \node[left=0pt of .west, scale=0.35] {$\text{g\_current}$};}
                                ]
                            ]
                            [,no edge, draw=none]
                        ]
                        [,no edge, draw=none]
                    ]
                    [,no edge, draw=none]
                ]
            ]
        ]
        [ 20, tikz={\node[right=0pt of .north east, scale=0.75]  {$(-\infty,\infty)$};}
            [15, tikz={\node[left=0pt of .north west, scale=0.75]  {$(-\infty,20)$};}]
            [30, tikz={\node[right=0pt of .north east, scale=0.75]  {$(20,\infty)$};}
                [25, tikz={\node[left=0pt of .north west, scale=0.75]  {$(20,30)$};}]
                [50, tikz={\node[right=0pt of .north east, scale=0.75]  {$(30,\infty)$};}
                    [45, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,50)$};}
                        [40, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,45)$};}
                            [35, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,40)$};}
                                [, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,35)$};}]
                                [k, tikz={\node[right=0pt of .north east, scale=0.75]  {$(35,40)$}; \node[left=0pt of .west, scale=0.35] {$\text{g\_current}$};}
                                    [, tikz={\node[left=0pt of .west, scale=0.75]  {$g_1$};}]
                                    [, tikz={\node[left=0pt of .west, scale=0.75]  {$g_2$};}]
                                ]
                            ]
                            [,no edge, draw=none]
                        ]
                        [,no edge, draw=none]
                    ]
                    [,no edge, draw=none]
                ]
            ]
        ]
    ]
\end{forest}

\begin{figure}[htb]
\centering
\includegraphics[width=150mm,scale=0.5]{FIG/extract_insert_2.png}
\caption{A visual depiction of the change in global state of BST during $\mathsf{insert(k,v)}$ operation, where key $k$ already exists in the tree $t$ }
\label{extract_insert2}
\end{figure}

\begin{forest}
for tree={
    grow=south,
    circle, draw, minimum size=3ex, inner sep=1pt,
    s sep=3mm
        }
    [, phantom, for children={fit=band}, s sep+=20mm,
         before drawing tree={%
          tikz+={%
            \node (a) [inner sep=0pt, fit=(!1) (!1L) (!1F)] {};
            \node (b) [inner sep=0pt, fit=(!l) (!lL) (!lF)] {};
            \node [anchor=center] at ($(a.east)!1/2!(b.west)$) {$\longrightarrow\quad$};
          },
        },
        [ 20, tikz={\node[right=0pt of .north east, scale=0.75]  {$(-\infty,\infty)$};}
            [15, tikz={\node[left=0pt of .north west, scale=0.75]  {$(-\infty,20)$};}]
            [30, tikz={\node[right=0pt of .north east, scale=0.75]  {$(20,\infty)$};}
                [25, tikz={\node[left=0pt of .north west, scale=0.75]  {$(20,30)$};}]
                [50, tikz={\node[right=0pt of .north east, scale=0.75]  {$(30,\infty)$};}
                    [45, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,50)$};}
                        [40, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,45)$};}
                            [35, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,40)$};}
                                [, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,35)$};}]
                                [, tikz={\node[right=0pt of .north east, scale=0.75]  {$(35,40)$}; \node[left=0pt of .west, scale=0.35] {$\text{g\_current}$};}
                                ]
                            ]
                            [$g_2$, draw=none]
                        ]
                        [,no edge, draw=none]
                    ]
                    [,no edge, draw=none]
                ]
            ]
        ]
        [ 20, tikz={\node[right=0pt of .north east, scale=0.75]  {$(-\infty,\infty)$};}
            [15, tikz={\node[left=0pt of .north west, scale=0.75]  {$(-\infty,20)$};}]
            [30, tikz={\node[right=0pt of .north east, scale=0.75]  {$(20,\infty)$};}
                [25, tikz={\node[left=0pt of .north west, scale=0.75]  {$(20,30)$};}]
                [50, tikz={\node[right=0pt of .north east, scale=0.75]  {$(30,\infty)$};}
                    [45, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,50)$};}
                        [40, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,45)$};}
                            [35, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,40)$};}
                                [, tikz={\node[left=0pt of .north west, scale=0.75]  {$(30,35)$};}]
                                [, tikz={\node[right=0pt of .north east, scale=0.75]  {$(35,40)$}; \node[left=0pt of .west, scale=0.35] {$\text{g\_current}$};}
                                ]
                            ]
                            [$g_2$, draw=none]
                        ]
                        [,no edge, draw=none]
                    ]
                    [,no edge, draw=none]
                ]
            ]
        ]
    ]
\end{forest}

@@
